### 2016-03-10 manuscript.Rnw
add paragraph about use of rotl for ecologists

The R programming language is a popular tool for phylogenetics and comparative
analysis. The R packages \texttt{ape} \citep{Paradis2004}, \texttt{phylobase}
\citep{phylobase-0.8.0}, and \texttt{RNeXML} \citep{Boettiger2015} each provide
functions to import and manipulate trees within R and save the results in
standard data formats. Additional packages including \texttt{phytools}
\citep{Revell2012}, \texttt{geiger} \citep{Pennell2014-geiger}, and
\texttt{ggtree} \citep{Yu-ggtree}, allow users to analyze and visualize data in
a phylogenetic context (see
\url{https://cran.r-project.org/web/views/Phylogenetics.html} for a
comprehensive list of phylogenetics packages in R). In addition to packages for
phylogenetic and comparative analyses, a growing number of R packages allows
users to query and access data from the web (e.g. \texttt{rFISHBASE}
\citep{Boettiger2012}, \texttt{rAvis} \citep{Varela2014}, and \texttt{paleoDB}
\citep{Varela2015}), such that data associated with taxa in a given phylogeny
can be obtained directly in R.  These packages, combined with the language's
support for literate programming (\citealt{Knuth1984}; \citealt{Xie2014}), make
R an ideal platform for reproducible research in phylogenetics and comparative
biology, as they allow a complete record of the steps taken in gathering,
processing, and analyzing a given data set to be produced.

>>>>

The R programming language is a popular tool for phylogenetics and comparative
analysis. The R packages \texttt{ape} \citep{Paradis2004}, \texttt{phylobase}
\citep{phylobase-0.8.0}, and \texttt{RNeXML} \citep{Boettiger2015} each provide
functions to import and manipulate trees within R and save the results in
standard data formats. Additional packages including \texttt{phytools}
\citep{Revell2012}, \texttt{geiger} \citep{Pennell2014-geiger}, and
\texttt{ggtree} \citep{Yu-ggtree}, allow users to analyze and visualize data in
a phylogenetic context (see
\url{https://cran.r-project.org/web/views/Phylogenetics.html} for a
comprehensive list of phylogenetics packages in R). In addition to packages for
phylogenetic and comparative analyses, a growing number of R packages allows
users to query and access data from the web (e.g. \texttt{rFISHBASE}
\citep{Boettiger2012}, \texttt{rAvis} \citep{Varela2014}, and \texttt{paleoDB}
\citep{Varela2015}), such that data associated with taxa in a given phylogeny
can be obtained directly in R.

In ecology, the development of the field of community phylogenetics
\citep{Webb2002} has created a need for researchers to have access to the
evolutionary relationships of species making up communities. The relative
contributions of the role of the environment (e.g., habitat filtering), and of
biotic interactions (e.g., competitive exclusion), are inferred from the
distribution of taxa on a phylogeny composed from species occurring at larger
spatial scale. R packages and other software have been developed to generate
phylogenies from species lists using taxonomic information or DNA
sequences. These phylogenies can then be used for community phylogenetics
analyses (e.g., \citep{Pearse2013}, \citep{Kembel}, \citep{Webb}), but as OTL
becomes more comprehensive, it could become a valuable resource for ecologists.

These packages, combined with the language's support for literate programming
(\citealt{Knuth1984}; \citealt{Xie2014}), make R an ideal platform for
reproducible research in phylogenetics and comparative biology, as they allow a
complete record of the steps taken in gathering, processing, and analyzing a
given data set to be produced.
